---
title: "An introduction to the 'LifeTable' package"
author: "Jason Thorpe [jdthorpe_at_gmail.com]"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{LifeTable}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r, echo=F}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Preliminaries

A life table divides a birth cohort into a discrete set of states such that 
for any given age, an individual in the cohort is place in exactly 1 state.
States in a two-state, might be "Alive" vs. "Dead", and in a three-state
life table might be "Alive without disease", "Alive with Disease", and
"Dead".  

The states that make up a Life tables can be represented by a graph such as:

```{r}
# make graphs here.
```

This package further assumes that the graph is acyclic, meaning that once an
individual moves out of a particular state, it is not possible to return to
that state.  In the three-state model, it is therefor not possible to return
to the "Alive without disease", after having departed that state to either
the "Alive with diseae" or "Dead" state. 

Creating an acyclig graph can often be achieved by defining states on the
events in an individual's history.  For example, we can re-define the three
state model by re-naming the categories "No Prior death or diagnosis",
"Prior Disease Diagnosis without Prior Death", and "Prior Death".  Whith
these lables, it is clear that even if an individual recovers from a disease
and is considered "disease free" from a clinical standpoint, that individual
still belongs in the second state in the state-space. 

Because the State-Space is acyclic, three will naturally be a source states
(states which only have arrows pointing away from them), zero or more
transient states (states with arrows pointing toward and away), and one or
more absorbing states (states which only have arrows pointing toward them.
In our three-state life table, "Alive without Disease" is a source state,
and "Dead" is an absorbing state.  When defining states based on the events
in an individual's history at a given age, there can be only one source state
("no prior events") but there can be multiple transient states (e.g. multiple
disease states) and absorbing states (e.g. Death With vs. without disease)

The *Incidence* of a given state is the number of individuals who arrive at
that state.  For example, when the three state model has run to completion
(the date of death of the last individual in the cohort), the incidence of
death will be 100%, the incidence of "disease" will between 0 and 100%, and
the incidence of healthy will be zero (since no-one 'arrives' at that
state).

When thinking about incidence, the 'States-as-a-set-of-prior-events' is
particularly useful, because we can define the incidence of a state as the
incidence of a particular event, such as disease diagnosis or death.  
For a given time interval $[t_0,t_1]$, the incidence depencs on the size of
the at-riak population and the risk of that particular event and the
duration of the time period (t_1 - t_0).  For source states, the incidence
during any period is zero, and for absorbing states, the
incidence is the difference between the portion of the population at that
state between $t_0$ and $t_1$.  For any given time period, the incidence of
a transient state will typically be greater than the difference between 
the number of individuals at that state between the start and end of the
time period.  For example, if the 'disease' state in the 3-state model is
highly lethal (Pancreatic cancer, for example), and the time periods were
relatively long (10 years), nearly all individuals who entered the disease
state during that interval would have died by the end of the interval.  


## Definitions

A *state table* is a matrix where each row represents the distribution of the
birth cohort across the state space at a specific age. 

A *discrete **net** hazard* ($h_{i,j}$) is the probability that an individual in
state $i$ will be in state $j$ at the end of a fixed period **in the absence of
all other competing risks**.  

A *continuous **net** hazard* $\lambda_{i,j}$ describes the rate that
individuals in state *i* transition to state *j* **in the absence of all other
competing risks**.   Continuous and discrete net hazards are related by the
equation 

 $h = 1 - e^{-\lambda \cdot (t_1 - t_0)}$

where $t_0$ and $t_1$ are the ages at the start and end of the defined
interval.   For convenience, the LifeTable packages assumes that 
$t_1 - t_0 = 1$, which simplifies the above relation to:

 $h = 1 - e^{-\lambda }$

A (continuous) transition marix ($\Lambda$) is an $N x N$ matrix where 
the $(i,j)^{th}$ element is the continuous net hazard ($\lambda_{i,j}$).

In the case when there is exactly one source (the first state in the
matrix, by convention), the transition matrix can be factored into the
absolute net hazards $\{\lambda_{i,1}\}$ and the *model matrix* $\{m_{i,j}\}$
with where 

\[ m_{i,j} = \left\{ \begin{array}{ll}
\dfrac{\lambda_{i,j}}{\lambda_{i,1}} & \mbox{if $\lambda_{i,1} \gt 0$};\\
0 & \mbox{otherwise}\end{array} \right. \]


Since the model matrix describes the relative risks of experieincing a
particular event conditional on the events in an individual's history, 
the model matrix will be typically be constructed from one or more Cox PH
models where each model is used to construct one row of the model matrix.

## Exercises

In this exercise, we will compose a life table to describe a cohort of women
who move through the following state space:

- Death without a prior diagnosis of Ovarian Cancer (Dead Without Cancer)
- Previous diagnosis with ovarian cancer (EOC)
- Ovaries Removed (with or without a previous tubal ligation or hysterectomy
  (Oophorectomized)
- Hysterecomized with or without prior tubal ligation (Hysterectomized)
- Tubes ligated (TL)
- None of the above (Healthy)

```{r}
# add figure here.
```

The first two states are *absorbing states* which together will eventually
absorb the entire population, and the last state on the 

## Creating Life Tables

To create a state table, we need (1) the initial state, (2) the Relative
Hazards matrix, and either (3.A) the incidence for each state or (3.B) the net hazards
for the reference state. 


The LifeTable package contains examples of state tables generated by the
[DevCan Software](http://surveillance.cancer.gov/devcan).  The dataset `EOC_develop` is a life table which includes 3
states: (1) Alive without Cancer, (2) Diagnosed with cancer, and (3) Dead
without cancer.


```{r}
library(LifeTable)
data(EOC_develop)
head(EOC_develop,3)
```

Since columns 5 and 6 in the `EOC_develop` contain the cumulative
probabilities of developing cancer or dying without cancer by the end of the
interval, we can create a state table with these 


```{r}
stateTable <- cbind(healthy = 1- EOC_develop$CumProbDevelopCancer 
							   - EOC_develop$CumProbDeathWithoutCancer,
		cancer = EOC_develop$CumProbDevelopCancer,
		DeathWithoutEOC = EOC_develop$CumProbDeathWithoutCancer) 
rownames(stateTable) <- seq(5,100,5)

# for completeness, we can add the state of the population at birth
stateTable <- rbind(
			birth = c( healthy = 1, # everyone starts out without cancer
					cancer = 0, 
					DeathWithoutEOC = 0) ,
			 stateTable)

head(stateTable,3)
```

As a trivial example, we can re-construct this life tables using the
`lifetable()` function using the incidence columns in `EOC_develop`. 


```{r}

# the hypothetical birth cohort of 10,000,000 begins healthy
initialState <- c(healthy=10000000, 
				  cancerOrDead=0,
				  DeathWithoutEOC=0)

# incidence numbers taken from EOC_develop 
incidenceMatrix <- cbind(
		healthy = 0,
		cancer = EOC_develop$DevelopCancer,
		DeathWithoutEOC = EOC_develop$DeathsWithoutCancer)
rownames(incidenceMatrix) <- rownames(EOC_develop)

# Hazard ratios are relative to the othere elements of in the row
modelMatrix <- matrix(
		c(0,0,0,
		  1,0,0,
		  1,0,0),nrow=3,byrow=T)

Cancer_Vs_Death <- lifeTable(
				  initialState=initialState,
				  incidence=incidenceMatrix,
				  hazardModel =modelMatrix )

head(Cancer_Vs_Death,3)

```

Notice that the numbers in above state table are relative to the size of our
model cohort of 10,000,000.  We can re-scale the above life table in order
to recapitulate the first example:

```{r}
head(Cancer_Vs_Death$states/10000000,3)
```

In addition to re-capitulating the state table, `lifeTable()` has also
calculated the net hazards for the referecne category ($\lambda_{i,1}$):


```{r}
head(Cancer_Vs_Death$netHazards,3)
```

## Expanding the life table

Transient states can be added to the model if either the
incidence or net hazards of each state additional state are knowon. 

Here, we'll add 3 transient states in order to incorporate the use of Tubal
Ligation, Hysterectomy with Salpingectomy (Removal of the uterus and tubes),
and Oophorectomy (removal of the ovaries) into our model.   The `LifeTable` 
package contains a dataset with the numbers of each of these procedures per 
100k persons in the United States in 2003.  

```{r}
data(GyneclogicalSureryRatesPer100k)

# (Notice that these data use very coarse age categories)
head(GyneclogicalSureryRatesPer100k,3)


# we're intereseted in all excision of the uterus, so we 
# need to include both sub-categories.
AnualHyterectomyRates <- (GyneclogicalSureryRatesPer100k$"Hysterectomy__abdominal_and_vaginal" +
					 GyneclogicalSureryRatesPer100k$"Other_excision_of_cervix_and_uterus") / 100000
AnualTubalLigationRates <- GyneclogicalSureryRatesPer100k$"Ligation_of_fallopian_tubes"/ 100000
AnualOophorectomyRates <- GyneclogicalSureryRatesPer100k$"Oophorectomy__unilateral_and_bilateral"/ 100000

```

The above rates need to be scaled to the US Female population
since we are modeling an all female cohort.   We can manage this conversion
using  2003 US census counts by sex included in the dataset
`US_population_2003`:


```{r}
data(US_population_2003)
head(US_population_2003,3)

US_population_2003[20,] <- US_population_2003[20,] + US_population_2003[21,]
US_population_2003 <- US_population_2003[-21,]

scale_factor <- US_population_2003$all / US_population_2003$female

AnualHyterectomyRates   <- AnualHyterectomyRates * scale_factor 
AnualTubalLigationRates <- AnualTubalLigationRates * scale_factor
AnualOophorectomyRates  <- AnualOophorectomyRates * scale_factor
```

Depending on how the numbers of procedures were surveyed, there are two ways we can
calculate incidence figures required to expand the life table.

### Method 1: Rates as Incidence

The denominator for the 'rate' variables calculated above is 
the estimated number alive of women in the observation period for each age
range. If these rates they can be used  calculate the incidence of each of 
the transient staes if by multiplying each by the proportion of the
original birth cohort that is alive during each age interval.  

However, the life table `EOC_develop`, does not have an 'alive' category,
and no combination of the three states captures the proportion alive wihtin 
each interval.  In order to calculate the incidence for our expanded life
table, we'll need to take advantage of another the dataset `EOC_death`
(also constructed by DevCan) which includes three states: (1) Alive (2) dead from
cancer, and (3) Dead without cancer.  

```{r}
data(EOC_death)
head(EOC_death,3)

# calculate the at-risk population * 
dead  <- EOC_death$"CumProbDeath" 
alive <- 1 - dead

# scale the rates to the proporition of the birth cohort
# that is alive in each age range rather than 100,000
HyterectomyIncidence   <- alive * 5 * AnualHyterectomyRates   / 100000 
TubalLigationIncidence <- alive * 5 * AnualTubalLigationRates / 100000 
OophorectomyIncidence  <- alive * 5 * AnualOophorectomyRates  / 100000 

# (Note that each age category spans 5 years, hence the factor of 5)

```

(* Technically this is the population size at the *end* of the interval, so for
this calculation, some kind of interpolation between the starting and ending
population sizes may be appropriate.)


### Method 2: Rates as Hazards

In the model bitth cohort, the at-risk population for the three transient states is
dwindling over time as women are diagnosed with cancer or die without
cancer.  However, this may not be true of the at-risk population used in
estimating the rates per 100k of each procedure.  In particular, the at-risk population 
that may be steady state (if younger women age into the at risk population
 at the approximatly the same rate that women die) or it may even be growing over 
time (due to variation in the birth and death rates over the last century). 
In the case of the steady-state at-risk population, the estimated rates can
be incuded as as net hazards ($\lambda$)for the alive population, and We can use
the relation 

 $h = 1 - e^{-\lambda * (t_1 - t_0)}$

to calculate the incidence as follows:

```{r}
# scale the rates to the proporition of the birth cohort
# that is alive in each age range rather than 100,000
HyterectomyIncidence_alt   <- alive * (1-exp( - 5 * AnualHyterectomyRates   / 100000 ))
TubalLigationIncidence_alt <- alive * (1-exp( - 5 * AnualTubalLigationRates / 100000 ))
OophorectomyIncidence_alt  <- alive * (1-exp( - 5 * AnualOophorectomyRates  / 100000 ))
```

In this instance, this adjustmemt makes little difference owing to the relatively low
rates of each event (since as $\lambda \rightarrow 0$, $\frac{1-\lambda}{e^{-\lambda}} 
\rightarrow 1 $ ).  


## Constructing the life table.  

To generate the state table, in addition to the incidence of each event at each
step, we need (1) an initial state vector and (2) the model matrix to
specify our model, as follows: 

```{r}

initialState <- c(healthy                = 1, # everyone starts out healthy 
				  TubesLigated           = 0, # incidence: Tubal Ligation
				  NoUterus               = 0, # incidence: hysterectomy
				  NoOvariesAndNoEOC      = 0, # incidence: oophorectomy 
				  OvarianCancer          = 0, # incidence: cancer
				  DeathWithoutEOC        = 0) # incidence: dead without cancer 

IncidenceTable <- cbind(
	healthy              = 0,
	TubesLigated         = TubalLigationIncidence_alt,
	NoUterus             = HyterectomyIncidence_alt,
	NoOvariesAndNoEOC    = OophorectomyIncidence_alt,
	OvarianCancer        = Cancer_Vs_Death$incidence[,'cancerOrDead']/10000000,
	DeathWithoutEOC      = Cancer_Vs_Death$incidence[,'DeathWithoutEOC']/10000000)

hazardModel <- matrix(
	c(  0,0,0,0,0,0,  # healthy              
		1,0,0,0,0,0,  # TubesLigated         
		1,1,0,0,0,0,  # NoUterus             
		1,1,1,0,0,0,  # NoOvariesAndNoEOC 
		1,1,1,0,0,0,  # OvarianCancer        
		1,1,1,1,0,0), # DeathWithoutEOC      
	nrow=6,
	byrow=T)

```

Each column in the model matrix indicates the states that are possible to
transition to *directly* from that state.  

- Column one indicates that individuals without a medical history are at
  risk for each of the events in the model.  
- Column two indicates that individuals with only a Tubal Ligation are at
  risk for any of the remaining events in the model.  
- Column three indicates that individuals with a Tubal Ligation and/or
  hysterectomy are at risk for Oophorectomy, EOC, and Death without EOC,
- Column four indicates that individuals with an oophorectomy (with or without
  prior Tubal Ligation or hysterectomy) are only at risk for Death without
  EOC. (It's hard to get ovarian cancer without ovaries.)
- Columns 5 and 6 indicate that Diagnosed with Ovarian Cancer and Death
  without Cancer  are absorbing states.  

Notice that our model has 6 states, and we have calculated 5 incidence
parameters.   This is because we have 5 states for which we will be
estimating hazard parameters, and one source state with a fixed hazard (zero)
which need not be estimated.  

We can now use `lifeTable()` to calculate a state table as follows:

```{r}

life_table_with_Procedures <- lifeTable(
		initialState=initialState,
		incidence   =IncidenceTable,
		hazardModel =hazardModel )

head(life_table_with_Procedures)

tail(life_table_with_Procedures,3)

```

## Relative Risks (Hazard Ratios).  

As mentioned earlier, the ratios of the elements within a row of the
model matrix define hazard ratios for the each event for individuals in states represented
by the respective columns. By convention, the reference category (no prior
events) is the first state, and hence the elements in first column of the
model matrix are set to 1 (with the exception of $m_{1,1}$ which is set to
zero).  However, the above model matrix `hazardModel`, all of the hazards
were equal to the reference hazard, which may not be true.

We could suppose that at least some hysterectomies are motivated (at least in
part) for sterilization, so perhaps women with an prior Tubal ligation would
be 20% less likely to undergo hysterectomy.  Furthermore, a subset of Ovarian
Cancers may arise in the tubes, so perhaps women with a tubal ligation are 30%
less likely to be diagnosed with Ovarian Cancer, and women with a hysterectomy
(in which fallopian tubes may be removed) could be at even lower risk of
Ovarian cancer (say 40%). [If you haven't guessed, each of the numbers in this
paragraph is made up.]  We specify this model as follows: 


```{r}

hazardModel_2 <- matrix(c(
		0,0    ,0    ,0,0,0,
		1,0    ,0    ,0,0,0,
		1,1-0.2,0    ,0,0,0, # rate/risk of hysterectomy is 20% lower in women with a prior TL
		1,1    ,1    ,0,0,0,
		1,1-0.3,1-0.4,0,0,0, # rate/risk of EOC is 30% lower in women with a TL and 40% lower in women with prior Hysterectomy.
		1,1    ,1    ,1,0,0),6,byrow=T)

life_table_with_Procedures <- lifeTable(
		initialState=initialState,
		incidence   =IncidenceTable,
		hazardModel =hazardModel )

head(life_table_with_Procedures,3)

tail(life_table_with_Procedures,3)

```

Notice that updating our hazard model affected the population states for the
transient states, but not the absorbing states, because this the absorbing
states can be calculated from the incidence alone.  However, updating the
hazard model *did* affect the net hazards for every outcome.

## Modeling an intervention

After calculating a life-table via `lifetable()` a population intervention
can be modeled by modifying some of the returned values, and creating a new 
life table.  For example, since Tubal Ligation may potentially prevent
Ovarian Cancer in some women, we may be interested in the effect of
conducting tubal ligation when child bearing is complete.  For simplicity,
we'll use age 40 in this example:


```{r}
# Extract the net Hazards matrix
netHazards_TL_at_40 <- life_table_with_Procedures$netHazards

# Set the net hazards of TL to infinity at age 40 which will
# result in all women having their tubes ligated at age 40
#  if they are in the at-risk population for TL at that time
netHazards_TL_at_40['40 <= Age < 45','NoUterus'] <- Inf

# Re-calculat the life table using the modified net hazard matrix
TL_at_40 <- lifeTable( initialState=initialState,
		netHazards  =netHazards_TL_at_40,
		hazardModel =hazardModel_2 )

tail(TL_at_40,3)
```

## What are my chances?

The above example shows the effect of tubal ligation at age 40 in a birth
cohort (for women who were in the at-risk population at 40).  However, for 
a woman who is age currently 40, the previous analysis does not directly address 
her chances for being diagnosed with Ovarian Cancer depending on whether 
*she* elects to have tubal ligation or not.  To address the question of what 
effect Tubal Ligation would have on *her* chances of being diagnosed with
EOC, we can create a pair of life tables beginning at age 40 which begin
with the reference population (n=1) having  TL or not:  


```{r}
# extract the net hazards matrix after age 40
Net_Hazards_after_40 <- life_table_with_Procedures$netHazards[-1:-8,]
head(Net_Hazards_after_40)

# a new initial state starting with the entire population (n=1)  
# in the 'tubes ligated' state.
initialState_TL <- c(healthy        = 0, # everyone starts out healthy 
				  TubesLigated      = 1, # incidence: Tubal Ligation
				  NoUterus          = 0, # incidence: hysterectomy
				  NoOvariesAndNoEOC = 0, # incidence: oophorectomy 
				  OvarianCancer     = 0, # incidence: cancer
				  DeathWithoutEOC   = 0) # incidence: dead without cancer 

# re-calculate the life table begining with TL at age 40
TL_at_age_40 <- lifeTable(
		initialState = initialState_TL,
		netHazards   = Net_Hazards_after_40,
		hazardModel  = hazardModel_2 )

tail(TL_at_age_40,3)

```

Similarly and the choice *not* to undergo TL can be modeled as follows: 

```{r}
# extract the net hazards matrix after age 40
Net_Hazards_after_40_No_TL <- Net_Hazards_after_40

# set the hazard of TL to zero, so that the woman will not ever undergo TL
Net_Hazards_after_40_No_TL[,'TubesLigated'] <- 0 

No_TL_at_age_40 <- lifeTable(
		initialState = initialState,
		netHazards   = Net_Hazards_after_40_No_TL,
		hazardModel  = hazardModel_2 )

tail(No_TL_at_age_40,3)
```

Looking at the state tables, we can see that the chances that our 40 year old
woman is diagnosed with EOC in her lifetime is 1.38% if she does not elect
to undergo Tubal ligation, and if she elects Tubal Ligation at age 40, her
chances drop to 0.97%.  (*Don't forget* that the hazard ratios in this
example were completely made up!)


## Expanding a life table by sub-dividing a state

Ovarian cancer has sub-types defined by histology which have substantially
different risks of death, which we may want to include in our life table model.
As always, to create a life table (which we'll use to calculate the net
hazards, we'll need (1) a start state, (2) a hazard model, and
(3) either incidence or net risks for each category.  We'll expand the
previous 6-state life table to include histology specific EOC, and the
easiest way to obtain the data required to expand the model is to separate
the incidence of EOC into sub-categories.

The table `HistologyByAge` contains the distributions of EOC subtypes by
age, and we can use this to sub-divide the incidence of EOC in our
life table (obtained from the SEER databases).  

```{r}
data(HistologyByAge)
head(HistologyByAge,3)

initialState <- c('healthy' = 1,
					 'TubesLigated' = 0,
					 'NoUterus' = 0,
					 'NoOvariesAndNoEOC' = 0,
					 'ClearCell' = 0,
					 'Endometrioid' = 0,
					 'Misc' = 0,
					 'Mucinous' = 0,
					 'Serous' = 0,
					 'DeathWithoutEOC ' = 0)

hazardModel <- matrix(
		c(  0,0    ,0    ,0,0,0,0,0,0,0,  # healthy              
			1,0    ,0    ,0,0,0,0,0,0,0,  # TubesLigated         
			1,1-0.2,0    ,0,0,0,0,0,0,0,  # NoUterus             
			1,1    ,1    ,0,0,0,0,0,0,0,  # NoOvariesAndNoEOC 
			1,1    ,1    ,0,0,0,0,0,0,0,  # OvarianCancer: ClearCell     
			1,1    ,1    ,0,0,0,0,0,0,0,  # OvarianCancer: Endometrioid         
			1,1    ,1    ,0,0,0,0,0,0,0,  # OvarianCancer: Misc                
			1,1    ,1    ,0,0,0,0,0,0,0,  # OvarianCancer: Mucinous            
			1,1-0.3,1-0.4,0,0,0,0,0,0,0,  # OvarianCancer: Serous              
			1,1    ,1    ,1,0,0,0,0,0,0), # DeathWithoutEOC      
		10,byrow=T)

hazardModel

# (Just re-asigning for clarity)
.LT = life_table_with_Procedures

IncidenceTable <- cbind(
		"healthy"           = .LT$incidence[,'healthy'],
		"TubesLigated"      = .LT$incidence[,'TubesLigated'],
		"NoUterus"          = .LT$incidence[,'NoUterus'],
		"NoOvariesAndNoEOC" = .LT$incidence[,'NoOvariesAndNoEOC'],
		'ClearCell'         = .LT$incidence[,'OvarianCancer'] * HistologyByAge$ClearCell,
		'Endometrioid'      = .LT$incidence[,'OvarianCancer'] * HistologyByAge$Endometrioid,
		'Misc'              = .LT$incidence[,'OvarianCancer'] * HistologyByAge$Misc,
		'Mucinous'          = .LT$incidence[,'OvarianCancer'] * HistologyByAge$Mucinous,
		'Serous'            = .LT$incidence[,'OvarianCancer'] * HistologyByAge$Serous,
		"DeathWithoutEOC "  = .LT$incidence[,'DeathWithoutEOC' ])

life_table_with_procedures_and_histology <- lifeTable(
		initialState=initialState,
		incidence   =IncidenceTable,
		hazardModel =hazardModel )

tail(life_table_with_procedures_and_histology,3)

```

Note that by subdividing the 'OvarianCancer' state into sub-states using the fraction
of the incidence, we preserve the total incidence by age.  This would not be
true if we subdivided the total hazard according to the proportions
of the incidence of the subtypes by age.  In fact, if we had done so, we
would have decreased the total incidence of EOC, and by the same token, by
preserving the total incidence we have increased the sum of the net hazards:

```{r}
all(apply(life_table_with_Procedures$netHazards,1,sum)<= 
	apply(life_table_with_procedures_and_histology$netHazards,1,sum))
```

This is expected because hazards are not additive.  Specifically because,
$1-e^{-(x + y)} < (1-e^{-x}) + (1-e^{-x})$ when $x>0$ and $y>0$.

## Generating PopSim Model Parameters


To parameterize the PopSim model, we need to generate cumulative (net)
distributions for each of the cancer subtypes and gynecologic procedures
(contained in the life table `life_table_with_procedures_and_histology` as well as death
from other causes (*not* `DeathWithoutEOC`), and we'll want to exclude the
distribution for `DeathWithoutEOC` so that it is not accidentally used in
the Model.  We'll first generate a simple life table object with alive vs.
dead, and then use the function `generateModelParams()` to generate the
parameters.  

Note that there is an additional argument to `generateModelParams()` called
`folder` which, if provided, will created two files *file/distributions.xml*
and *folder/DecisionRule.py* with the distribution and Decision Rule files,
respectively.


```{r,eval=FALSE}
generateModelParams <- function(ages=seq(0,100,5),
							life_table_with_procedures_and_histology,
							what to do here?
							use dot args to pass multiple lifetables?
							exclude='DeathWithoutEOC')
decisionRuleWriter <- function(life_table_with_procedures_and_histology)
```

Note that the above code is *almost* everything you need to implement the
PopSim model for a whole new disease!


## Methodology

This is where a description of the methodology will go.



